---
title: "Adult"
author: "TM"
date: "2016. február 28."
output: html_document
---
```{r, warning = FALSE, message = FALSE, echo = FALSE}
library(dplyr)
library(ggplot2)
library(randomForest)
library(data.table)
library(pander)
library(knitr)
opts_knit$set(root.dir = "../")
options(stringsAsFactors = FALSE)

##### Opening and basic formating of datasets


#setwd("D:/Google Drive/Mikike/BusinessAnalytics/Tananyag/Data Science for Business/Final project")

train.adult.df <- read.csv("adult.data", header = FALSE, sep = ",")
colnames(train.adult.df) <- c("age","workclass","fnlwgt","education","education_num","marital_status","occupation","relationship","race","sex","capital_gain","capital_loss","hours_per_week","native_country","earning")
setDT(train.adult.df)
train.adult.df[train.adult.df==" ?",]=NA

test.adult.df <- read.csv("adult.test", header = FALSE, sep = ",")
colnames(test.adult.df) <- c("age","workclass","fnlwgt","education","education_num","marital_status","occupation","relationship","race","sex","capital_gain","capital_loss","hours_per_week","native_country","earning")
setDT(test.adult.df)
test.adult.df[test.adult.df==" ?",]=NA
test.adult.df <- test.adult.df[-1,]

dt <- train.adult.df


```
#Executive summary




This **Documentation** was created to introduce a process of classifying  `r nrow(dt)` adults by their earnings based on their social and professional parameters.There are two earning groups set up: >50k and <=50k.

There are **two basic dataset** for the analysis. One for training the models and one for testing. Both datasets have the same structure. The aim of analysis is finding a model that can predict the earning category of an individual based on the characteristics (features) provided. Characteristics (features):

   **Feature: type**

- age: continuous
- workclass: category
- fnlwgt: continuous
- education: category
- education-num: continuous
- marital-status: category
- occupation: category
- relationship: category
- race: category
- sex: category
- capital-gain: continuous
- capital-loss: continuous
- hours-per-week: continuous
- native-country

**Process model building**, first the test and training sets are compared if they follow the same pattern. Second step, features are checked if they are suitable for the analysis. Third, based on the findings features are modified. As a fourts step, models are built and evaluated. Fifth, step models are tested on theseparated test set to verify if they provide good results on a separated dataset as well.

**Dataset credits**: Lichman, M. (2013). UCI Machine Learning Repository: http://archive.ics.uci.edu/ml. Irvine, CA: University of California, School of Information and Computer Science.

#Testing variables & features


When loading datasets all non-numeric variables are treated as characters as in case of basic statistics factors sometimes mix up the calculations.


### Target variable: earning

Target variable is a binary variable with two possible values:
- >50k , meaning yearly earnings above USD 50,000 and 
- <=50k , meaning yearly earnings below (or equal to) USD 50,000.

***Characteristics of earnings***
```{r, echo=FALSE}
pander(summary (dt$earning))
```

***Number of observations in each group***
```{r, echo=FALSE}
pander(table(dt$earning))
```

### Training set vs Test set


Test and training set is to be compared if they show the same distribution. As they are provided as separate datasets, it should be tested if they have the same characteristics, really randomly separated ones.

```{r, echo=FALSE, warning=FALSE}

##### Training set vs Test set
#class(train.adult.df)
templist <- list(train.adult.df[, type := "train"], test.adult.df[, type := "test"])
dtCompare <- rbindlist(templist, use.names=TRUE)

for (i in c(2,4,6,7,8,9,10, 14,15)) {
  dtCompare[[i]]<-as.factor(dtCompare[[i]])
}
dtCompare[[1]]<-as.numeric(dtCompare[[1]])
#class(dtCompare$workclass)

ggplot(dtCompare, aes(x=age))+geom_histogram(binwidth = 1)+facet_grid(type~., scales = "free")
ggplot(dtCompare, aes(x=workclass))+geom_bar()+facet_grid(type~., scales = "free")+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.45))
ggplot(dtCompare, aes(x=fnlwgt))+geom_bar(binwidth = 50000)+facet_grid(type~., scales = "free")
ggplot(dtCompare, aes(x=education))+geom_bar()+facet_grid(type~., scales = "free")+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.45))
ggplot(dtCompare, aes(x=education_num))+geom_bar()+facet_grid(type~., scales = "free")
ggplot(dtCompare, aes(x=marital_status))+geom_bar()+facet_grid(type~., scales = "free")+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.45))
ggplot(dtCompare, aes(x=occupation))+geom_bar()+facet_grid(type~., scales = "free")+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.45))
ggplot(dtCompare, aes(x=relationship))+geom_bar()+facet_grid(type~., scales = "free")+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.45))
ggplot(dtCompare, aes(x=race))+geom_bar()+facet_grid(type~., scales = "free")
ggplot(dtCompare, aes(x=sex))+geom_bar()+facet_grid(type~., scales = "free")
ggplot(dtCompare, aes(x=capital_gain))+geom_bar(binwidth = 10000)+facet_grid(type~., scales = "free")
ggplot(dtCompare, aes(x=capital_loss))+geom_bar(binwidth = 1000)+facet_grid(type~., scales = "free")
ggplot(dtCompare, aes(x=hours_per_week))+geom_bar()+facet_grid(type~., scales = "free")
ggplot(dtCompare, aes(x=native_country))+geom_bar()+facet_grid(type~., scales = "free")+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.45))

```

Test and training sets show the same distribution, so I decided to use them in this division.

### Checking features - training set


All the features should be checked if they are suitable for using in model building. I prepare distribution plots, check for missing values and outliers. 

#### Feature: age

```{r, echo=FALSE, warning=FALSE}
plot(dt$age, main = "Age distribution", ylab = "Age", ylim = c(0, 110))
ggplot(dt, aes(x=age))+geom_histogram(binwidth = 1)+facet_grid(earning~., scales = "free")
```

***General statistics of age***
```{r, echo=FALSE, warning=FALSE}
pander(summary(dt$age))
```

The number of mising values is `r sum(is.na(dt$age))`. As it could be seen the number of 90-year-old participants is quite high compared to the neighbouring values. The reason for this could be from more sources, e.g. it really covers all ages >=90. 

```{r, echo=FALSE, warning=FALSE}
# strange 90-year-old population - remove 90s
a <- NULL
for(i in 17:90){
  a <- c(a, paste(i, ":",dt[age==i, .N]))
}
```


***Distribution of age, in numbers***

```{r, echo=FALSE, warning=FALSE}
pander(a)

dt[, logage:= log(age)]
#ggplot(dt, aes(x=logage))+geom_histogram(binwidth = .02)+facet_grid(earning~., scales = "free")
```

The number of these observations (in total) is not high, so I decided to drop age 90 as outlying numbers.

 
 
#### Feature: workclass

The number of missing values is quite high, `r sum(is.na(dt$workclass))`, in percents `r sum(is.na(dt$workclass))/length(dt$workclass)*100`%. This is quite high, it should be decided to tranform or drop these observations.

```{r, echo=FALSE, warning=FALSE}
dt$workclass = as.factor(dt$workclass)
ggplot(dt, aes(x=workclass))+geom_bar()+facet_grid(earning~., scales = "free")+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.45))
```

```{r, echo=FALSE, warning=FALSE}
#unique(dt$workclass)
dt$workclass = as.character(dt$workclass)
# many NAs
```

#### Feature: fnlwgt

This feature is a generated weight number based on social, economical and professional charactristics of the persons. Number of missing values is `r sum(is.na(dt$fnlwgt))`. General statistics show that the distribution is skewed:

```{r, echo=FALSE, warning=FALSE}
summary(dt$fnlwgt)
```

that could be followed on the histogram as well:
```{r, echo=FALSE, warning=FALSE}
ggplot(dt, aes(x=fnlwgt))+geom_histogram(binwidth = 30000)+facet_grid(earning~., scales = "free")
```

Number of values of *fnlwgt* above 750,000 is `r dt[fnlwgt>750000, .N]`. Although that is a low number I decided to keep every observation, as a result of eyeballing the chart below:

```{r, echo=FALSE, warning=FALSE}
plot(dt$fnlwgt, main = "fnlwgt distribution", ylab = "fnlwgt")
dt[, logfnlwgt:= log(fnlwgt)]
#ggplot(dt, aes(x=logfnlwgt))+geom_histogram(bins = 50)
```



#### Feature: education

The number of missing values is `r sum(is.na(dt$education))`, with a minimal skew in distribution. 

```{r, echo=FALSE, warning=FALSE}
dt$education = as.factor(dt$education)
ggplot(dt, aes(x=education))+geom_bar()+facet_grid(earning~., scales = "free")+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.45))
#unique(dt$education)

dt$education = as.character(dt$education)
```


####Feature: education_num

The number of missing values is `r sum(is.na(dt$education_num))`. General statistics:

```{r, echo=FALSE, warning=FALSE}
pander(summary(dt$education_num))
```



```{r, echo=FALSE, warning=FALSE}
#plot(dt$education_num, main = "education_num distribution", ylab = "education_num")
ggplot(dt, aes(x=education_num))+geom_histogram(binwidth = 1)+facet_grid(earning~., scales = "free")
```



####Feature: marital status


The number of missing values is `r sum(is.na(dt$marital_status))`. 

```{r, echo=FALSE, warning=FALSE}
dt$marital_status = as.factor(dt$marital_status)
ggplot(dt, aes(x=marital_status))+geom_bar()+facet_grid(earning~., scales = "free")+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.45))
#unique(dt$marital_status)

dt$marital_status = as.character(dt$marital_status) 
```


####Feature: occupation


The number of missing values is quite high, `r sum(is.na(dt$occupation))`, in percents `r sum(is.na(dt$occupation))/length(dt$occupation)*100`%. This is quite high, it should be decided to tranform or drop these observations.

```{r, echo=FALSE, warning=FALSE}
dt$occupation = as.factor(dt$occupation)
ggplot(dt, aes(x=occupation))+geom_bar()+facet_grid(earning~., scales = "free")+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.45))
#unique(dt$occupation)
dt$occupation = as.character(dt$occupation) 
```


####Feature: relationship


The number of missing values is `r sum(is.na(dt$relationship))`. 

```{r, echo=FALSE, warning=FALSE}
dt$relationship = as.factor(dt$relationship)
ggplot(dt, aes(x=relationship))+geom_bar()+facet_grid(earning~., scales = "free")+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.45))
#unique(dt$relationship)
dt$relationship = as.character(dt$relationship)
```


####Feature: race


The number of missing values is `r sum(is.na(dt$race))`. 

```{r, echo=FALSE, warning=FALSE}
dt$race = as.factor(dt$race)
ggplot(dt, aes(x=race))+geom_bar()+facet_grid(earning~., scales = "free")
#unique(dt$race)
dt$race = as.character(dt$race)
```


####Feature: sex


The number of missing values is `r sum(is.na(dt$sex))`. 

```{r, echo=FALSE, warning=FALSE}
dt$sex = as.factor(dt$sex)
ggplot(dt, aes(x=sex))+geom_bar()+facet_grid(earning~., scales = "free")
#unique(dt$sex)
dt$sex = as.character(dt$sex)
```


####Feature: capital_gain


Number of missing values is `r sum(is.na(dt$capital_gain))`. General statistics:


```{r, echo=FALSE, warning=FALSE}
pander(summary(dt$capital_gain))
```


Plots:

```{r, echo=FALSE, warning=FALSE, message=FALSE}
plot(dt$capital_gain, main = "capital_gain distribution", ylab = "capital_gain")
ggplot(dt, aes(x=capital_gain))+geom_histogram()+facet_grid(earning~., scales = "free")

```

Plot for capital gain > 0

```{r, echo=FALSE, warning=FALSE, message=FALSE}
kingsOfTheHill <- dt[capital_gain>0,]
ggplot(kingsOfTheHill, aes(x=capital_gain))+geom_histogram()+facet_grid(earning~., scales = "free")
```

Number of obsevations where capital gain > 0
```{r, echo=FALSE, warning=FALSE, message=FALSE}
kingsOfTheHill[,.N]
```

According to the plots and general statistics capital gain of 99,999 looks suspicious. Probably, the same situation as at 'age'. I consider these errorneous data.




####Feature: capital_loss


Number of missing values is `r sum(is.na(dt$capital_loss))`. General statistics:

```{r, echo=FALSE, warning=FALSE, message=FALSE}
pander(summary(dt$capital_loss))
```

Plots:

```{r, echo=FALSE, warning=FALSE, message=FALSE}
plot(dt$capital_loss, main = "capital_loss distribution", ylab = "capital_loss")
ggplot(dt, aes(x=capital_loss))+geom_histogram()+facet_grid(earning~., scales = "free")
```

Plot for capital loss > 0
```{r, echo=FALSE, warning=FALSE, message=FALSE}
kingsOfTheDeep <- dt[capital_loss>0,]
ggplot(kingsOfTheDeep, aes(x=capital_loss))+geom_histogram()+facet_grid(earning~., scales = "free")
```

Number of obsevations where capital loss > 0
```{r, echo=FALSE, warning=FALSE, message=FALSE}
kingsOfTheDeep[,.N]
```

According to the plots there is no need for adjusting this feature.


####Feature: hours_per_week

Number of missing values is `r sum(is.na(dt$hours_per_week))`. General statistics:

```{r, echo=FALSE, warning=FALSE}
pander(summary(dt$hours_per_week))
```

Distribution plots
```{r, echo=FALSE, warning=FALSE}
# 100 hours? - 84 and above - drop?
plot(dt$hours_per_week, main = "hours_per_week distribution", ylab = "hours_per_week")
ggplot(dt, aes(x=hours_per_week))+geom_histogram(binwidth = 1)+facet_grid(earning~., scales = "free")
```

```{r, echo=FALSE, warning=FALSE}
a <- NULL
for(i in 1:99){
  a <- c(a, paste(i, ":",dt[hours_per_week==i, .N]))
}

pander(a)

```

Working hours above 84 per week make no sense - daily 12 hours, 7 days, and the plots and numbers shows that there are some problems with these figures. I plan to drop data above 84 hours/week.



####Feature: Native country

Number of missing values is `r sum(is.na(dt$native_country))`, in percentage `r sum(is.na(dt$native_country))/length(dt$native_country)*100`. 

```{r, echo=FALSE, warning=FALSE}
dt$native_country = as.factor(dt$native_country)
ggplot(dt, aes(x=native_country))+geom_bar()+facet_grid(earning~., scales = "free")+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.45))
#unique(dt$native_country)
```

Percentage of native countries (nr of persons):
```{r, echo=FALSE, warning=FALSE}
country <- data.table(table(dt$native_country, useNA = "always"))
colnames(country) <- c("Country", "Nr")
a <-  sum(country$Nr)
country[, Percent :=  round(Nr/a*100,4)]
pander(head(country[with(country,order(-Percent)),], 10))
dt$native_country = as.character(dt$native_country)
```

##### Checking workclass and occupation NA-s

Workclass and Occupation have very similar number of NAs. The recors with NAs are the same in `r sum(is.na(dt$occupation)==is.na(dt$workclass))/length(dt$workclass)`%. 

#Feature transformation


Decisions:
- introducing new category for workclass and occupation NAs: UNKNOWN both in test and training/evaluation sets,
- removing all remaining NAs from training/evaluation set,
- removing 90-year-old observations.
- removing capital gain of 99,999,
- removing weekly working hous of 84 or more.

```{r, echo=FALSE, warning=FALSE}
training_base <- dt[,-18, with = FALSE]
training_base <- training_base[,-17, with = FALSE]
training_base <- training_base[,-16, with = FALSE]

#fill missing variables in Workclass and occupation features
test.adult.df$workclass[is.na(test.adult.df$workclass)] = "UNKNOWN"
test.adult.df$occupation[is.na(test.adult.df$occupation)] = "UNKNOWN"
training_base$workclass[is.na(training_base$workclass)] = "UNKNOWN"
training_base$occupation[is.na(training_base$occupation)] = "UNKNOWN"

# Deleting observations with NAs
training_base <- na.omit(training_base)
test.adult.df <- na.omit(test.adult.df)

# Correcting Test set target and columns
setDT(test.adult.df)
test.adult.df[earning==" <=50K.", earning := " <=50K"] 
test.adult.df[earning==" >50K.", earning := " >50K"] 
test.adult.df <- test.adult.df[,-16, with = FALSE]

# removing 90 year-olds 
training_base <- filter(training_base, training_base$age!=90)
test.adult.df <- filter(test.adult.df, test.adult.df$age!=90)  

# removing hours_per_week > 83 (daily 12)
training_base <- filter(training_base, training_base$hours_per_week<83)
test.adult.df <- filter(test.adult.df, test.adult.df$hours_per_week<83)
```

Transformations above means 3% data loss.



#Model building and selection


After the modifications above I divided the training dataset into a training and a validation set. 

I chose **Random Forest** and **GBM** methods.I used H20 for procesing.
set.seed(73)


```{r, echo=FALSE, warning=FALSE}
set.seed(73)
N <- nrow(training_base)
id_tr <- sample(1:N,2*N/3)
id_va <- sample(base::setdiff(1:N, id_tr))

# chk sum(id_tr)+sum(id_va)-(32561+1)*32561/2

trainingSet <- training_base[id_tr,]
validationSet <- training_base[id_va,]


### H2O # http://localhost:54321

library(h2o)
h2o.init(max_mem_size = "1g", nthreads = -1)

#paste(names(trainingSet), collapse = ",")
for (i in c(2,4,6,7,8,9,10, 14,15)) {
  trainingSet[[i]]<-as.factor(trainingSet[[i]])
  validationSet[[i]]<-as.factor(validationSet[[i]])
  test.adult.df[[i]]<-as.factor(test.adult.df[[i]])
}
test.adult.df[[1]]<-as.numeric(test.adult.df[[1]])

#str(trainingSet)
#str(test.adult.df)
#unique(trainingSet$earning)  
#unique(validationSet$earning)
#unique(test.adult.df$earning)

### models
#class(test.adult.df$earning)

dtTr <- as.h2o(trainingSet)
dtTr$earning <- as.factor(dtTr$earning)
dtVa <- as.h2o(validationSet)
dtVa$earning <- as.factor(dtVa$earning)
dtTe <- as.h2o(test.adult.df)
dtTe$earning <- as.factor(dtTe$earning)
```


# gmb
```{r, echo=FALSE}

system.time({
  md <- h2o.gbm(x = seq(ncol(dtTr) - 1), y = ncol(dtTr), 
                training_frame = dtTr, validation_frame = dtVa,
                max_depth = 15, ntrees = 500, learn_rate = 0.01, nbins = 200,
                stopping_rounds = 3, stopping_tolerance = 1e-3)
})
md

h2o.auc(md)
h2o.auc(h2o.performance(md, dtTe))
```





```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

