---
title: "Adult"
author: "TM"
date: "2016. február 28."
output: html_document
---
```{r, echo=FALSE}
library(dplyr)
library(ggplot2)
library(randomForest)
library(data.table)

library(knitr)
opts_knit$set(root.dir = "../")
options(stringsAsFactors = FALSE)
##### Opening and basic formating of datasets


#setwd("D:/Google Drive/Mikike/BusinessAnalytics/Tananyag/Data Science for Business/Final project")
opts_knit$set(root.dir = "../")
train.adult.df <- read.csv("adult.data", header = FALSE, sep = ",")
colnames(train.adult.df) <- c("age","workclass","fnlwgt","education","education_num","marital_status","occupation","relationship","race","sex","capital_gain","capital_loss","hours_per_week","native_country","earning")
setDT(train.adult.df)
train.adult.df[train.adult.df==" ?",]=NA

test.adult.df <- read.csv("adult.test", header = FALSE, sep = ",")
colnames(test.adult.df) <- c("age","workclass","fnlwgt","education","education_num","marital_status","occupation","relationship","race","sex","capital_gain","capital_loss","hours_per_week","native_country","earning")
setDT(test.adult.df)
test.adult.df[test.adult.df==" ?",]=NA
test.adult.df <- test.adult.df[-1,]


dt <- train.adult.df


```

This Markdown 'r nrow(dt)' was made as to document a machine learning process, that aims a classification of  a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
